---
title: "Download HBV parameters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Download-HBV-parameters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(HBVr)
library(tmap)
library(terra)
library(sf)
```


## 1) Disclaimer & Preprocessing logic
Beneath is the example how can one download and preprocess HBV parameter maps created by Beck et al. ([2020](http://www.gloh2o.org/hbv/)). The initial dataset was distributed as NetCDF file. It contains ten cross-validation folds of 12 parameters:

- **BETA** -- Shape coefficient of recharge function

- **FC** -- Maximum soil moisture storage (mm)

- **K0** -- Recession coefficient of upper zone (day<sup>−1</sup>)

- **K1** -- Recession coefficient of upper zone (day<sup>−1</sup>)

- **K2** -- Recession coefficient of lower zone (day<sup>−1</sup>)

- **LP** -- Soil moisture value above which actual evaporation reaches potential evaporation

- **PERC** -- Maximum percolation to lower zone (mm day<sup>−1</sup>)

- **UZL** -- Threshold parameter for extra outflow from upper zone (mm)

- **TT** -- Threshold temperature (°C)

- **CFMAX** -- Degree‐day factor (mm °<sup>−1</sup>1 day<sup>−1</sup>)

- **CFR** -- Refreezing coefficient

- **CWH** -- Water holding capacity

We converted every fold to the Cloud-Optimized GeoTiff (COG) using the following GDAL command: `gdal_translate ordinary.tif COG.tiff -co COPY_SRC_OVERVIEWS=YES -co COMPRESS=LZW -co TILED=YES`. The final COG was validated using the python [script](https://github.com/OSGeo/gdal/blob/master/swig/python/gdal-utils/osgeo_utils/samples/validate_cloud_optimized_geotiff.py). Thus we ended up with 10 COGs with 12 bands each.

<!-- Overall logic of the preprocessing workflow can be visualized as follows: -->


## 2) Example workflow

The essential aim of the `hbv_get_parameters` function is to ease the accessibility of HBV parameter maps created by Beck et al. ([2020](http://www.gloh2o.org/hbv/)). Therefore this function allows to download any of the cross-validation folds for your particular area of interest (AOI).

Let's imagine, you need to download **first**, **second** and **fifth** folds for the Luxembourg.

```{r, message=TRUE, warning=FALSE, out.width = "90%", fig.align='center'}
library(HBVr)
library(tmap) # for vizualisztion purposes
library(sf) # for vizualisztion purposes

# Locate the shapefile
f <- system.file("ex/lux.shp", package="terra")
# Read it as SpatVector
v <- vect(f)

# Plot it!
tmap_mode("view")

tm_basemap("Stamen.Terrain") +
  tm_shape(st_as_sf(v)) +
  tm_polygons()
```

Initially, the `hbv_get_parameters` function has two types of output. First is the mean zonal statistics which is computed with `terra::global` in the background. Second one is the simple `SpatRaster` output.

To download 1, 2 and 5 folds we need to specify it in the function. Here is the example how can we retrieve mean values for the entire Luxembourg.

```{r, message=FALSE, warning=TRUE}
zonal_stat <- 
  hbv_get_parameters(
    aoi = v,
    folds = c(1, 2, 5),
    mean = TRUE
  )
```

The output of `zonal_stat` is hidden below ↓ 
<!-- <div class="fold s o"> -->
```{r}
zonal_stat
```
<!-- </div> -->


To download `SpatRaster` objects we need to run the following code chunk:
```{r, message=FALSE, warning=TRUE}

rasters <- 
  hbv_get_parameters(
    aoi = v,
    folds = c(1, 2, 5),
    mean = FALSE
  )

```

The output of `rasters` is hidden below ↓ 
<!-- <div class="fold s o"> -->
```{r}
rasters
```
<!-- </div> -->

The function returns a list of `SpatRaster` objects. The first one, i.e. Fold 1, can be plotted like this:

```{r message=FALSE, warning=FALSE, out.width="50%"}
plot(rasters[[1]])
```

It can be plotted with `tmap` for example:
```{r message=FALSE, warning=FALSE, out.width="100%"}
# Plot it!
tmap_mode("view")

tm_basemap("Stamen.Terrain") +
  tm_shape(rasters[[1]]$beta) +
  tm_raster()
```
